CFHTLens Weak Lensing Emulator and Cosmological Constraints from the Minkowski Functionals and Moments

### 2014-11-05 Paper_minkowski/paper.tex
added ray tracing simulations section

\begin{equation}
\label{costfunction}
d(\mathcal{D}) = \frac{2D^{1/2}}{N(N-1)}\sum_{i<j}^N\frac{1}{\vert\mathbf{x}_i-\mathbf{x}_j\vert}
\end{equation} 
%
This problem is mathematically equivalent to the minimization of the Coulomb potential energy of $N$ unit charges in a unit box, which will make sure that the charges are as evenly spread as possible throughout the confining volume. Finding the optimal design $\mathcal{D}_m$ that minimized (\ref{costfunction}) can be computationally very demanding, and hence we decided to use a simplified approach that, although approximate, serves our purposes for the grid design. We use an iterative procedure:
\begin{enumerate}
\item We start from the diagonal design $\mathcal{D}_0$: $x_i^d\equiv i/(N-1)$
\item We shuffle the coordinates of the particles in each dimension independently $x_i^d = \mathcal{P}_d\left(\frac{1}{N-1},\frac{2}{N-1},...,1\right)$ where $\mathcal{P}_1,...,\mathcal{P}_D$ are random independent permutations of $(1,2,...,N)$
\item We pick a random particle pair $(i,j)$ and a random coordinate $d\in\{1,...,D\}$ and swap $x_i^d\leftrightarrow x_j^d$
\item We compute the new cost function, if this is less than the previous step, we keep the exchange, otherwise we revert the coordinate swap
\item We repeat steps 3 and 4 until the relative cost function change is less than a chosen accuracy parameter $\epsilon$ 
\end{enumerate}
%
We found that for $N=91$ grid points, order of $10^5$ iterations are sufficient to reach an accuracy of $\epsilon\sim10^{-4}$; once the optimal design $\mathcal{D}_m$ has been found, we can invert the mapping $\Pi\rightarrow[0,1]^3$ to find our simulation parameter sampling $\mathbf{p}_s$, which we show in 
Figure \ref{designfig}.
%
\begin{figure*}
\begin{center}
\includegraphics[scale=0.4]{Figures/design.eps}
\caption{$(\Omega_m,w)$ and $(w,\sigma_8)$ projections of our the simulation design; the blue points correspond to the CFHTemu1 simulation set, which consists of one $N$--body simulation per point, while the red point corresponds to the CFHTcov simulation set, which is based on 50 independent $N$--body simulations}
\label{designfig}
\end{center}
\end{figure*}
%
For each parameter point on the grid $\mathbf{p}_s$ we run one $N$--body simulation and perform ray tracing through it, as described in \S~\ref{raysim}, to simulate CFHTLens shear catalogs; this set of simulations will be called CFHTemu1 throughout the rest of this work. Additionally, we run 50 independent $N$--body simulations with a \textit{fiducial} parameter choice $\mathbf{p}_0=(0.26,-1.0,0.8)$ for the purpose of measuring accurately the covariance matrices which will serve the parameter inferences as in \S-\ref{cosmostats}; this set of simulations will be called CFHTcov throughout the rest of this work.    
\subsection{Ray Tracing Simulations}
\label{raysim}

>>>>

\begin{equation}
\label{costfunction}
d(\mathcal{D}) = \frac{2D^{1/2}}{N(N-1)}\sum_{i<j}^N\frac{1}{\vert\mathbf{x}_i-\mathbf{x}_j\vert}
\end{equation} 
%
This problem is mathematically equivalent to the minimization of the Coulomb potential energy of $N$ unit charges in a unit box, which will make sure that the charges are as evenly spread as possible throughout the confining volume. Finding the optimal design $\mathcal{D}_m$ that minimized (\ref{costfunction}) can be computationally very demanding, and hence we decided to use a simplified approach that, although approximate, serves our purposes for the grid design. We use an iterative procedure:
\begin{enumerate}
\item We start from the diagonal design $\mathcal{D}_0$: $x_i^d\equiv i/(N-1)$
\item We shuffle the coordinates of the particles in each dimension independently $x_i^d = \mathcal{P}_d\left(\frac{1}{N-1},\frac{2}{N-1},...,1\right)$ where $\mathcal{P}_1,...,\mathcal{P}_D$ are random independent permutations of $(1,2,...,N)$
\item We pick a random particle pair $(i,j)$ and a random coordinate $d\in\{1,...,D\}$ and swap $x_i^d\leftrightarrow x_j^d$
\item We compute the new cost function, if this is less than the previous step, we keep the exchange, otherwise we revert the coordinate swap
\item We repeat steps 3 and 4 until the relative cost function change is less than a chosen accuracy parameter $\epsilon$ 
\end{enumerate}
%
We found that for $N=91$ grid points, order of $10^5$ iterations are sufficient to reach an accuracy of $\epsilon\sim10^{-4}$; once the optimal design $\mathcal{D}_m$ has been found, we can invert the mapping $\Pi\rightarrow[0,1]^3$ to find our simulation parameter sampling $\mathbf{p}_s$, which we show in 
Figure \ref{designfig}.
%
\begin{figure*}
\begin{center}
\includegraphics[scale=0.4]{Figures/design.eps}
\caption{$(\Omega_m,w)$ and $(w,\sigma_8)$ projections of our the simulation design; the blue points correspond to the CFHTemu1 simulation set, which consists of one $N$--body simulation per point, while the red point corresponds to the CFHTcov simulation set, which is based on 50 independent $N$--body simulations}
\label{designfig}
\end{center}
\end{figure*}
%
For each parameter point on the grid $\mathbf{p}_s$ we run one $N$--body simulation and perform ray tracing through it, as described in \S~\ref{raysim}, to simulate CFHTLens shear catalogs; this set of simulations will be called CFHTemu1 throughout the rest of this work. Additionally, we run 50 independent $N$--body simulations with a \textit{fiducial} parameter choice $\mathbf{p}_0=(0.26,-1.0,0.8)$ for the purpose of measuring accurately the covariance matrices which will serve the parameter inferences as in \S-\ref{cosmostats}; this set of simulations will be called CFHTcov throughout the rest of this work.

\subsection{Ray Tracing Simulations}
\label{raysim}
The goal of this paragraph is to give an outline of our simulation pipeline; the fluctuations in the dark matter density field between a source at redshift $z$ and an observer located on Earth will cause small deflections to the trajectories of light rays travelling from the source to the observer. The fluctuations in the dark matter field are described by their gravitational potential $\Phi(\mathbf{x},z)=\Phi(\mathbf{x}_\perp,w(z))$, where we can trade the physical coordinates $\mathrm{x}$ with the comoving distance from the observer $w$ and two transverse coordinates $\mathbf{x}_\perp=w\pmb{\beta}$ using the flat sky approximation. Here $\pmb{\beta}$ refers to the angular coordinate on the sky of a physical point $\mathbf{x}$, as seen from the observer. We estimate the dark matter gravitational potential running $N$--body simulations (with $N=512^3$) with the public code Gadget2 \citep{Gadget2}, using a comoving box size of $240h^{-1}$Mpc. Using a similar procedure as in \citep{RayTracingJain,RayTracingHartlap}, the equation that governs the light ray deflections can be written in the form
\begin{equation}
\label{raytrajectory}
\frac{d^2\mathbf{x}(w)}{dw^2} = -\frac{2}{c^2}\nabla_{\mathbf{x}_\perp}\Phi(\mathbf{x}_\perp(w),w)
\end{equation}
%
where $\mathbf{x}(w)$ is the trajectory of a single light ray. Suppose that a light ray reaches the observer at an angular position $\pmb{\theta}$ on the sky: we want to know where this light ray originated, knowing it comes from a redshift $z_s$. To answer this question we need to integrate equation (\ref{raytrajectory}) with the initial conditions $\pmb{\beta}(0;\pmb{\theta})=\pmb{\theta}$, $\dot{\pmb{\beta}}(0;\pmb{\theta})=0$ up to a distance $w_s=w(z_s)$ to obtain the source angular position $\pmb{\beta}(w_s;\pmb{\theta})$; for the light ray trajectory solver, based on equation (\ref{raytrajectory}), we use our proprietary implementation Inspector Gadget. Once we know the details of the light ray trajectories, we can easily infer the weak lensing interesting quantities by taking the angular derivatives of the ray deflections $A(w_s;\pmb{\theta}) = \partial \pmb{\beta}(w_s;\pmb{\theta})/\partial\pmb{\theta}$ and performing the usual spin decomposition to infer the convergence $\kappa$ and the shear components $(\gamma_1,\gamma_2)$
%
\begin{equation}
A(w_s;\pmb{\theta}) = (1-\kappa(w_s;\pmb{\theta}))\pmb{I} - \gamma_1(w_s;\pmb{\theta})\sigma_3 - \gamma_2(w_s;\pmb{\theta})\sigma_1
\end{equation}  
%
where $\pmb{I}$ is the $2\times2$ identity and $\sigma_{1,3}$ are the first and third Pauli matrices. $\kappa$ is related to the source apparent magnification, while $(\gamma_1,\gamma_2)$ are related to the source apparent ellipticity, as seen from the observer.
